# Gaussian Process Function
```{r}
# generate data
library(ggplot2)
library(magrittr)
library(tidyverse)
gaussprocess <- function(from = 0,
                         to = 1,
                         K = function(s, t) {min(s, t)},
                         start = NULL,
                         m = n) {
    t <- seq(from = from, to = to, length.out = m)
    Sigma <- sapply(t, function(s1) {
        sapply(t, function(s2) {
            K(s1, s2)
        })
    })
    path <- MASS::mvrnorm(mu = rep(0, times = m), Sigma = Sigma)
    if (!is.null(start)) {
        path <- path - path[1] + start  # Must always start at "start"
    }
    return(data.frame("t" = t, "xt" = path))
}
```
# Generate Simulation Data
```{r}
# 
N = 20
n = 100
t = seq(0,1,length=n)

set.seed(1000)
## Gaussian Process
gaussian_data_generate = function(times = N) {
    gaussian_data <- gaussprocess(K = function(s, t) {exp(-16 * (s - t) ^ 2)})
    for(i in 1:(times-1)){
        temp <- gaussprocess(K = function(s, t) {exp(-16 * (s - t) ^ 2)})[,2]
        gaussian_data <- cbind(gaussian_data, temp)
    }
    return(gaussian_data[,-1])
}

gaussian_data = gaussian_data_generate(times = N)
matplot(gaussian_data,type = "l")

data_generate = function(n,N,t){
  beta = t^3
  x = c()
  y = c()
  for (i in 1:N) {
    x_temp = exp(sin(t))+rnorm(n,0,0.1)
    y_temp = beta*x_temp+gaussian_data[,i]*x_temp+rnorm(n,0,0.5)
    x = cbind(x,x_temp)
    y = cbind(y,y_temp)
  }
  return(list(x = x,y = y))
}

data = data_generate(n,N,t)
x = data$x
y = data$y
matplot(t,y,type = "l")

```


# Estimate Model, REML-Based Backfitting Algorithm
```{r}
# kernel function
ptm = proc.time()

library(lava)
# Gaussian Kernel
kh_fun=function(tdiff,h){
  u=tdiff/h
  kh=1/((2*pi)^0.5)*exp(-(u^2)/2)/h
  return(kh)
}


h = 0.015

L = 0.8
tfit = seq(0,1,length = n)

sigma2.t =  c()
bhatlist = vector("list",n)
alphahat = c()
for (j in 1:n) {
  tt = tfit[j]
  tdiff = tfit-tt
  hi_tt = cbind(rep(1,n),tdiff)
  
  Xilist = vector("list",N)
  Xmat = c()
  for (i in 1:N) {
    Xilist[[i]] = cbind(x[,i]*hi_tt[,1],x[,i]*hi_tt[,2])
    Xmat = rbind(Xilist[[i]])
  }
  
  # Initial Values of Variance Components
  D = diag(c(1,1))
  sigma2 = 1; Ini = diag(rep(1,n))
  Ri = sigma2*Ini
  
  Khi = diag(kh_fun(tdiff,h))
  
  epsilon = 100
  while (epsilon>L) {
    Vilist = vector("list",N)
    omigahilist = vector("list",N)
    for (i in 1:N) {
      Vilist[[i]] = Ri + Khi^0.5%*%Xilist[[i]]%*%D%*%t(Xilist[[i]])%*%Khi^0.5
      omigahilist[[i]] = Khi^0.5%*%solve(Vilist[[i]])%*%Khi^0.5
    }
    
    alphajhat.term1 = 0
    alphajhat.term2 = 0
    for (i in 1:N) {
      alphajhat.term1 = alphajhat.term1 + t(Xilist[[i]]) %*% omigahilist[[i]] %*% Xilist[[i]]
      alphajhat.term2 = alphajhat.term2 + t(Xilist[[i]]) %*% omigahilist[[i]] %*% y[,i]
    }
    alphajhat = solve(alphajhat.term1) %*% alphajhat.term2
    
    bjhat = c()
    for (i in 1:N) {
      bijhat = D %*% t(Xilist[[i]]) %*% omigahilist[[i]] %*% (y[,i]-Xilist[[i]]%*%alphajhat)
      bjhat = cbind(bjhat,bijhat)
    }
    
    # Estimating Variance Components
    ehat = c()
    Pvi_term = 0
    for (i in 1:N) {
      ehat_i = Khi^0.5 %*% (y[,i] - Xilist[[i]] %*% alphajhat - Xilist[[i]] %*% bjhat[,i])  
      ehat = cbind(ehat,ehat_i)
      Pvi_term = Pvi_term + t(Xilist[[i]]) %*% omigahilist[[i]] %*% Xilist[[i]]
    }
    
    Pvilist = vector("list",N)
    for (i in 1:N) {
      Pvilist[[i]] = solve(Vilist[[i]])-solve(Vilist[[i]])%*%Khi^0.5%*%Xilist[[i]]%*%solve(Pvi_term)%*%t(Xilist[[i]])%*%Khi^0.5%*%solve(Vilist[[i]])
    }
  
    sigma2_next = 0
    D_next = 0
    for (i in 1:N) {
      sigma2_next = sigma2_next + t(ehat[,i]) %*% ehat[,i] + sigma2*(n-sigma2*tr(Pvilist[[i]]))
      D_next = D_next + bjhat[,i]%*%t(bjhat[,i])+
        (D-D%*%t(Xilist[[i]])%*%Khi^0.5%*%Pvilist[[i]]%*%Khi^0.5%*%Xilist[[i]]%*%D)
    }
    sigma2_next = (sigma2_next/(N*n))
    D_next = (D_next/n)
    
    # Termination Criteria for Variance Component Estimation
    numerator = sum(abs(as.vector(D_next)-as.vector(D)))
    denominator = sum(abs(as.vector(D)))
    
    numerator = abs(norm(D_next,"2")-norm(D,"2"))
    denominator = abs(norm(D,"2"))
    
    epsilon = numerator/denominator
    D = (diag(diag(D_next)))
    sigma2 = sigma2_next[1,1]
  }
  sigma2.t = c(sigma2.t,sigma2)
  alphahat = cbind(alphahat, alphajhat)
  bhatlist[[j]] = bjhat
  print(j)
}

(proc.time()-ptm)/60
```


```{r}
# End of Point Estimation Procedure; Extract Estimated Parameters
betahat = alphahat[1,]
vihatlist = c()
for (i in 1:n) {
  vihatlist = rbind(vihatlist,bhatlist[[i]][1,])
}
yfit = c()
for (i in 1:N) {
  yfit = cbind(yfit,x[,i]*betahat + x[,i]*vihatlist[,i])
}


matplot(t,y,type = "l",ylim = c(-5.5,7),lwd = 1)
for (i in 1:N) {
  lines(tfit,yfit[,i],type = "l",col = "red", lwd = 1.5)
}


i = 1
plot(t,y[,i],type = "l",ylim = c(-8,8))
lines(tfit,yfit[,i],type = "l",col = "red", lwd = 1.5)
```
# Extract Parameters - Covariance Matrix
```{r}
# Calculate Matrix Form of r(s, t)
r2mat = abs(cov(t(vihatlist)))
```

# Randomly Generate a New Curve
```{r}
set.seed(2)
library(MASS)
data_new = data_generate(n,N=1,t)
x_new = data_new$x
y_new = data_new$y
tao1 = 30
tao2 = 70
delta1 = 5
delta2 = 10
y_new[(tao1+1):(tao2)] = y_new[(tao1+1):(tao2)]+ x_new[(tao1+1):(tao2)]*delta1
y_new[(tao2+1):n] = y_new[(tao2+1):n]+ x_new[(tao2+1):n]*delta2
plot(y_new,type = "l")
```

# Calculate the covariance matrix, transform the data of t1 to tm, and compute the statistics.
```{r}
Statistic.vcm = c()

for (m in 1:n) {

  xm = x_new[1:m,]
  tm = t[1:m]
  ym = y_new[1:m,]-xm*betahat[1:m]

  Rm = diag(sigma2.t[1:m])
  if (m == 1) {
    SIGMA_m = xm *r2mat[1:m,1:m] * xm + sigma2.t[1:m]
  }else{
    SIGMA_m = diag(xm) %*% diag(diag(r2mat[1:m,1:m])) %*% t(diag(xm)) +Rm
  }

  xm_tilde = ginv(SIGMA_m^0.5) %*% xm
  ym_tilde = ginv(SIGMA_m^0.5) %*% ym
  
  lambda = 0.8
  wjlist = c()
  deltahat = c()
  for (j in 1:m) {
    wj = (1-lambda)^(m-j)
    wjlist = c(wjlist,wj)
    deltahat = c(deltahat, sum(xm_tilde[1:j]*ym_tilde[1:j]*wj)/sum(xm_tilde[1:j]*xm_tilde[1:j]*wj))
  }
  ym_tilde.fit = xm_tilde * deltahat

  RSS0 = t(ym_tilde)  %*% diag(wjlist) %*% (ym_tilde)
  RSS1 = t(ym_tilde-ym_tilde.fit) %*% diag(wjlist)  %*% (ym_tilde-ym_tilde.fit)
  temp = (RSS0-RSS1)
  Statistic.vcm = c(Statistic.vcm,temp)
}
plot(Statistic.vcm,type = "o",cex=0.5)
```

# Recursive computation of statistic
```{r}
Hm = 0
Hlist = c()
for (mm in 1:n) {
  xm = x_new[1:mm,]
  tm = t[1:mm]
  ym = y_new[1:mm,]-xm*betahat[1:mm]

  Rm = diag(sigma2.t[1:mm])
  if (mm == 1) {
    SIGMA_m = xm *r2mat[1:mm,1:mm] * xm + sigma2.t[1:mm]
  }else{
    SIGMA_m = diag(xm) %*% diag(diag(r2mat[1:mm,1:mm])) %*% t(diag(xm)) +Rm
  }

  xm_tilde = ginv(SIGMA_m^0.5) %*% xm
  ym_tilde = ginv(SIGMA_m^0.5) %*% ym
  
  w = 1-lambda
  wjlist = c()
  deltahat = c()
  for (j in 1:mm) {
    wj = (1-lambda)^(mm-j)
    wjlist = c(wjlist,wj)
    deltahat = c(deltahat, sum(xm_tilde[1:j]*ym_tilde[1:j]*wj)/sum(xm_tilde[1:j]*xm_tilde[1:j]*wj))
  }
  
  Hm.next = w*Hm-(xm_tilde[mm]%*%deltahat[mm])^2+2*ym_tilde[mm]*t(xm_tilde[mm])%*%deltahat[mm]
  Hm = Hm.next
  Hlist = c(Hlist,Hm)
}
plot(Hlist,type = "o",cex=0.5)

```
# Expectation of the statistic
```{r}
EH = 0
EHlist = c()
for (mm in 1:n) {
  xm = x_new[1:mm,]
  tm = t[1:mm]
  ym = y_new[1:mm,]-xm*betahat[1:mm]

  Rm = diag(sigma2.t[1:mm])
  if (mm == 1) {
    SIGMA_m = xm *r2mat[1:mm,1:mm] * xm + sigma2.t[1:mm]
  }else{
    SIGMA_m = diag(xm) %*% diag(diag(r2mat[1:mm,1:mm])) %*% t(diag(xm)) +Rm
  }

  xm_tilde = ginv(SIGMA_m^0.5) %*% xm
  ym_tilde = ginv(SIGMA_m^0.5) %*% ym
  
  w = 1-lambda
  wjlist = c()
  for (j in 1:mm) {
    wj = (1-lambda)^(mm-j)
    wjlist = c(wjlist,wj)
  }

  if (mm == 1) {
    Gmm = solve(xm_tilde[mm]%*%t(xm_tilde[mm]))
  }else{
    Gmm = solve(w*solve(Gm)+xm_tilde[mm]%*%t(xm_tilde[mm]))
  }
  # Gmm = solve(sum(xm_tilde*(xm_tilde)*wjlist[1:mm]))
  
  term1 = w*EH
  term2 = 0
  for (j in 1:mm) {
    term2 = term2 + wjlist[j]^2*(t(xm_tilde[mm])%*%Gmm%*%xm_tilde[j])^2
  }
  term3 = 2*t(xm_tilde[mm])%*%Gmm%*%xm_tilde[mm]
  
  EH.next = term1-term2+term3
  EH = EH.next
  Gm = Gmm
  EHlist = c(EHlist,EH)
}
plot(EHlist,type = "o",cex=0.5)
```

# Recursive formula for the variance of the statistic E_UUH
```{r}
Gm = 0 
E2UHm = 0 
E2UHmlist = c()
E2UHmlist = c(E2UHmlist,E2UHm)
epsilon4 = 3
for (mm in 2:n) {
  xm = x_new[1:mm,]
  tm = t[1:mm]
  ym = y_new[1:mm,]-xm*betahat[1:mm]

  Rm = diag(sigma2.t[1:mm])
  if (mm == 1) {
    SIGMA_m = xm *r2mat[1:mm,1:mm] * xm + sigma2.t[1:mm]
  }else{
    SIGMA_m = diag(xm) %*% diag(diag(r2mat[1:mm,1:mm])) %*% t(diag(xm)) +Rm
  }

  xm_tilde = ginv(SIGMA_m^0.5) %*% xm
  ym_tilde = ginv(SIGMA_m^0.5) %*% ym
  
  w = 1-lambda
  m = mm-1
  # Gm
  if (m == 1) {
    Gmm = solve(xm_tilde[m]%*%t(xm_tilde[m]))
  }else{
    Gmm = solve(w*solve(Gm)+xm_tilde[m]%*%t(xm_tilde[m]))
  }
  Gm = Gmm
  # term1
  term1 = w * E2UHm
  
  # term2
  # term2.1
  term2.1 = 0
  for (j in 1:m) {
    wj = w^(m-j)
    qmj = t(xm_tilde[m])%*%Gm%*%xm_tilde[j]
    term2.1 = term2.1 + wj^4*xm_tilde[j]%*%t(xm_tilde[j])%*%qmj^2*epsilon4
  }
  # term2.2
  m_c = c(1:m)
  term2.2 = 0
  term2.3 = 0
  if (m == 1) {
    term2.2 = 0
    term2.3 = 0
  }else{
    for (j in m_c) {
      qmj = t(xm_tilde[m]) %*% Gm %*% xm_tilde[j]
      wj = w^(m-j)
      for (i in m_c[-j]) {
        wi = w^(m-i)
        qmi = t(xm_tilde[m]) %*% Gm %*% xm_tilde[i]
        term2.2 = term2.2 + wj^2 * wi^2 * xm_tilde[j] %*% t(xm_tilde[j]) * qmi^2
        term2.3 = term2.3 + wj^2 * wi^2 * xm_tilde[j] %*% t(xm_tilde[i]) * qmi * qmj
      }
    }
  }
  term2 = term2.1 + term2.2 + term2.3
  # term3
  term3.1 = 0
  term3.2 = 0
  term3.3 = 0
  Pm_1 = 0
  if (m == 1) {
    term3.1 = 0
    term3.2 = 0
    term3.3 = 0
    Pm_1 = 0
  }else{
    for (j in 1:(m-1)) {
      wj = w^(m-j)
      qmj = t(xm_tilde[m]) %*% Gm %*% xm_tilde[j]
      term3.1 = term3.1+w^2*wj^2*xm_tilde[j]%*%qmj%*%xm_tilde[j]
      Pm_1 = Pm_1 + wj^2*xm_tilde[j]%*%t(xm_tilde[j])
    }
    term3.2 = w^2*t(xm_tilde[m])%*%Gm%*%xm_tilde[m]%*%Pm_1
    term3.3 = w^2*(xm_tilde[m])%*%t(xm_tilde[m])%*%Gm%*%Pm_1
  }
  term3.4 = xm_tilde[m]%*%t(xm_tilde)[m]*Gm%*% xm_tilde[m]%*%t(xm_tilde[m])*epsilon4
  term3 = 2*(term3.1 + term3.2 + term3.3 + term3.4)
  # term4
  if (m == 1) {
    term4.1 = 0
  }else{
    term4.1 = w*EHlist[m-1]
  }
  term4.2 = 0
  for (j in 1:(m)) {
    wj = w^(m-j)
    qmj = t(xm_tilde[m]) %*% Gm %*% xm_tilde[j]
    term4.2 = term4.2+wj^2*qmj^2
  }
  term4.3 = 2*Gm%*%(xm_tilde[m])
  term4 = term4.1+term4.2+term4.3
  
  E2UHmm = w^2*(term1-term2+term3) + xm_tilde[mm]%*%xm_tilde[mm]*term4
  E2UHm = E2UHmm
  E2UHmlist = c(E2UHmlist,E2UHm)
}
E2UHmlist
plot(E2UHmlist,type = "o",cex=0.5)
```
# Recursive formula for the covariance between A and B.
```{r}
cov_AB = c()
Gm = 0
for (mm in 1:n) {
  xm = x_new[1:mm,]
  tm = t[1:mm]

  Rm = diag(sigma2.t[1:mm])
  if (mm == 1) {
    SIGMA_m = xm *r2mat[1:mm,1:mm] * xm + sigma2.t[1:mm]
  }else{
    SIGMA_m = diag(xm) %*% diag(diag(r2mat[1:mm,1:mm])) %*% t(diag(xm)) +Rm
  }

  xm_tilde = ginv(SIGMA_m^0.5) %*% xm
  
  if (mm == 1) {
    Gmm = solve(xm_tilde[mm]%*%t(xm_tilde[mm]))
    EHm = 0
  }else{
    Gmm = solve(w*solve(Gm)+xm_tilde[mm]%*%t(xm_tilde[mm]))
    EHm = EHlist[mm-1]
  }
  w = 1-lambda
  term1 = w*t(xm_tilde[mm])%*%Gmm*E2UHmlist[mm]%*%Gmm%*%xm_tilde[mm]
  
  term2.1 = 0
  for (j in 1:mm) {
    wj = w^(mm-j)
    qmj = t(xm_tilde[mm]) %*% Gmm %*% xm_tilde[j]
    term2.1 = term2.1 + wj^2*qmj^2
  }
  
  term2 = w*term2.1*EHm
  
  cov_AB = c(cov_AB,term1-term2)
  Gm = Gmm
}
plot(cov_AB,type = "o",cex=0.5)
```
# varB
```{r}
Gm = 0
varB = c()
w = 1-lambda
for (mm in 1:n) {
  xm = x_new[1:mm,]
  tm = t[1:mm]

  Rm = diag(sigma2.t[1:mm])
  if (mm == 1) {
    SIGMA_m = xm *r2mat[1:mm,1:mm] * xm + sigma2.t[1:mm]
  }else{
    SIGMA_m = diag(xm) %*% diag(diag(r2mat[1:mm,1:mm])) %*% t(diag(xm)) +Rm
  }
  
  xm_tilde = ginv(SIGMA_m^0.5) %*% xm
  
  if (mm == 1) {
    Gmm = solve(xm_tilde[mm]%*%t(xm_tilde[mm]))
  }else{
    Gmm = solve(w*solve(Gm)+xm_tilde[mm]%*%t(xm_tilde[mm]))
  }
  
  
  term1 = 0
  for (j in 1:mm) {
    wj = w^(mm-j)
    qmj = t(xm_tilde[mm]) %*% Gmm %*% xm_tilde[j]
    term1 = term1 + wj^4*qmj^4*(epsilon4-1)
  }
  
  m_c = c(1:mm)
  term2 = 0
  if (mm == 1) {
    term2 = 0
  }else{
    for (j in m_c) {
      wj = w^(mm-j)
      qmj = t(xm_tilde[mm]) %*% Gmm %*% xm_tilde[j]
      for (i in m_c[-j]) {
        wi = w^(mm-i)
        qmi = t(xm_tilde[mm]) %*% Gmm %*% xm_tilde[i]
        term2 = term2 + wj^2 * wi^2 * qmi^2 * qmj^2
      }
    }
  }
  varB = c(varB,term1+term2)
  Gm = Gmm
}
plot(varB,type = "o",cex=0.5)
```
# varC
```{r}
Gm = 0
varC = c()
w = 1-lambda
for (mm in 1:n) {
  xm = x_new[1:mm,]
  tm = t[1:mm]

  Rm = diag(sigma2.t[1:mm])
  if (mm == 1) {
    SIGMA_m = xm *r2mat[1:mm,1:mm] * xm + sigma2.t[1:mm]
  }else{
    SIGMA_m = diag(xm) %*% diag(diag(r2mat[1:mm,1:mm])) %*% t(diag(xm)) +Rm
  }
  
  xm_tilde = ginv(SIGMA_m^0.5) %*% xm
  
  if (mm == 1) {
    Gmm = solve(xm_tilde[mm]%*%t(xm_tilde[mm]))
  }else{
    Gmm = solve(w*solve(Gm)+xm_tilde[mm]%*%t(xm_tilde[mm]))
  }
  
  
  term1 = 0
  m = mm-1
  if (m == 0) {
    term1 = 0
  }else{
    for (j in 1:m) {
      wj = w^(m-j)
      qmj = t(xm_tilde[mm]) %*% Gmm %*% xm_tilde[j]
      term1 = term1 + 4*w^2*wj^2*qmj^2
    }
  }
  term2 = 4*(t(xm_tilde[mm]) %*% Gmm %*% xm_tilde[mm])^2*(epsilon4-1)
  
  varC = c(varC,term1+term2)
  Gm = Gmm
}
plot(varC,type = "o",cex=0.5)
```

# cov_BC
```{r}
Gm = 0
cov_BC = c()
w = 1-lambda
for (mm in 1:n) {
  xm = x_new[1:mm,]
  tm = t[1:mm]


  Rm = diag(sigma2.t[1:mm])
  if (mm == 1) {
    SIGMA_m = xm *r2mat[1:mm,1:mm] * xm + sigma2.t[1:mm]
  }else{
    SIGMA_m = diag(xm) %*% diag(diag(r2mat[1:mm,1:mm])) %*% t(diag(xm)) +Rm
  }
  
  xm_tilde = ginv(SIGMA_m^0.5) %*% xm
  
  if (mm == 1) {
    Gmm = solve(xm_tilde[mm]%*%t(xm_tilde[mm]))
  }else{
    Gmm = solve(w*solve(Gm)+xm_tilde[mm]%*%t(xm_tilde[mm]))
  }
  
  
  term1 = (t(xm_tilde[mm]) %*% Gmm %*% xm_tilde[mm])^3*(epsilon4-1)
  
  term2 = 0
  m = mm-1
  if (m == 0) {
    term2 = 0
  }else{
    for (j in 1:m) {
      wj = w^(m-j)
      qmj = t(xm_tilde[mm]) %*% Gmm %*% xm_tilde[j]
      term2 = term2 + 2*w^2*(t(xm_tilde[mm]) %*% Gmm %*% xm_tilde[m])*wj^2*qmj^2
    }
  }
  
  cov_BC = c(cov_BC,2*(term1+term2))
  Gm = Gmm
}
plot(cov_BC,type = "o",cex=0.5)
```

# Vm
```{r}
cov_AC = 0
Vmlist = c()
Vm = 0
for (j in 1:mm) {
  Vm_next = w^2*Vm+varB[j]+varC[j]-2*cov_AB[j]+2*cov_AC-2*cov_BC[j]
  Vm = Vm_next
  Vmlist = c(Vmlist,Vm)
}
plot(Vmlist,type = "o",cex=0.5)
```
# Calculation of the standardized statistic
```{r}
HHlist = (Hlist-EHlist)/Vmlist^0.5
plot(HHlist[2:100],type = "o",cex=0.5)
plot(Hlist,type = "o",cex=0.5)
```


